---
title: "Confound it all!"
author: "Lab 3B"
date: 'Directions: Follow along with the slides, completing the questions in <font color=#25679E><b>blue</b></font> on your computer, and answering the questions in **red** in your journal.  
  
  <br>
  <br>
  Space, Click, Right Arrow or swipe left to move to the next slide.'
output: 
  slidy_presentation: 
    css: ../../IDSLabCSSRev.css
    mathjax: ../../extras/mathjax-local/MathJax.js
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

```{r, eval=TRUE, echo=FALSE, results='hide', message=FALSE, warning=FALSE}
library(mobilizr)
```

## Finding data in new places


- Since your first forays into doing data science, you've used data from two sources:
    - Built-in datasets from RStudio.
    - Campaign data from the Campaign Manager.
- Data can be found in many other places though, especially online.
- In this lab, we'll read an _observational study_ dataset from a website.
    - We'll use this data to then explore what factors are associated with a person's lung capacity.


## Importing our data

- Rather than _export_-ing the data and then _upload_-ing and _importing_-ing it, we'll pull the data straight from the webpage into R.

- You can find the data online here:
    - (Right-click and select _Open in New Window_) [https://raw.githubusercontent.com/thinkdataed/dataset/main/fev.csv](https://raw.githubusercontent.com/ThinkDataEd/dataset/main/fev.csv)
    
- <span class="blue">Click on the _Import Dataset_ button under the _Environment_ tab.</span>
    - Then click on the _From Text (readr)_ option.
    - Type or copy/paste the URL into the box.
    - Click _Update_.
- <span class="blue">Before importing, change the following _Import Options_:</span>
    - Name: `lungs`
    - _Uncheck_ the _First Row as Names_  
    - Change _Delimiter_ to _Whitespace_

## Our new data

- Variables that were measured include:
    - Age in years.
    - Lung capacity, measured in liters.
    - The youth's heights, in inches 
    - Genders; `"1"` for males, `"0"` for females.  
    - Whether the participant was a smoker, `"1"`, or non-smoker `"0"`.

## About the data

- The data come from the _Forced Expiratory Volume (FEV)_ study that took place in the late 1970's.
- The observations come from a sample of 654 youths, aged 3 to 19, in/around East Boston.
- Researchers were interested in answering the _research question_: 

_What is the effect of childhood smoking on lung health?_


## Cleaning your data

- <span class="blue">Now that we've got the data loaded, we need to clean it to get it ready for use _(look at lab 1F for help)_. Specifically:</span>
    - We want to name the variables: `"age"`, `"lung_cap"`, `"height"`, `"gender"`, `"smoker"`, in that order.
    - **(1) Write and run code changing the type of variable for `gender` and `smoker` from _numeric_ to _character_.**
- <span class="blue">After changing the variable types for `gender` and `smoker`:</span>
    - **(2) For `gender`, write and run code using `recode` to change `"1"` to `"Male"` and `"0"` to `"Female"`.**
    - **(3) For `smoker`, write and run code using `recode` to change `"1"` to `"Yes"` and `"0"` to `"No"`.**
    

## Analyzing our data

- Our `lungs` data is from an observational study.
- **(4) Write down a reason the researchers couldn't use an experiment to test the effects of smoking on children's lungs.**
- Observational studies are often helpful for analyzing how variables are related.
- **(5) Do you think that a person's age affects their lung capacity? Make a sketch of what you think a scatterplot of the two variables would look like and explain.**
- **(6) Write and run code using the `lungs` data to create an `xyplot` of `age` and `lung_cap`. Interpret the plot and describe why the relationship between the two variables makes sense.**

    
## Smoking and lung capacity

- **(7) Write and run code making a plot that can be used to answer the statistical investigative question:**

_Do people who smoke tend to have lower lung capacity than those who do not smoke?_

- **(8) Use your plot to answer the question**.
    - **(9) Were you surprised by the answer? Why?**
    - **(10) Can you suggest a possible confounding factor that might be affecting the result?**


## Let's compare

- **(11) Write and run code creating three subsets of the data:**
    - one that includes _only_ 13-year-olds ...
    - one that includes _only_ 15-year-olds ...
    - and one that includes _only_ 17-year-olds.
- **(12) Write and run code making a plot that compares the lung capacity of smokers and non-smokers for each subset.**
- **(13) How does the relationship between smoking and lung capacity change as we increase the age from 13 to 15 to 17?**


## Sum it up!

- **(14) Does smoking affect lung capacity? If so, how?**
    - Support your answers with appropriate plots. 
    - Explain why you included the variables you used in your plots.