---
title: "Random Sampling"
author: "Lab 3C"
date: 'Directions: Follow along with the slides, completing the questions in <font color=#25679E><b>blue</b></font> on your computer, and answering the questions in **red** in your journal.  
  
  <br>
  <br>
  Space, Click, Right Arrow or swipe left to move to the next slide.'
output: 
  slidy_presentation: 
    css: ../../IDSLabCSSRev.css
    mathjax: ../../extras/mathjax-local/MathJax.js
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

```{r, eval=TRUE, echo=FALSE, results='hide', message=FALSE, warning=FALSE}
library(mobilizr)
```

## Learning by sampling

- In many circumstances, there's simply no feasible way to gather data about everyone in a _population_.
    - For example, the Department of Water & Power (DWP) wants to determine how much water people in Los Angeles use to take a shower. They've created a survey to pass out to collect this information.
    - **(1) Write down two reasons why getting _everyone_ in Los Angeles to fill out the survey would be difficult. Also, write a sentence why the DWP might consider using a sample of households instead.**
    
- In this lab, we'll learn how _sampling methods_ affect how _representative_ a sample is of a _population_.


## Loading a population

- In previous labs, we used the `cdc` data as a sample for young people in the United States. 
    - In this lab, we'll consider these survey respondents to be our population.
- **(2) Load the `cdc` data into `R` and fill in the blanks to take a _convenience_ sample of the first 50 people in the data:**

```{r, eval = FALSE, echo = TRUE}
s1 <- slice(____, 1:____)
```

- **(3) Why do you think we call this method a _convenience_ sample?**


## Comparing your convenience sample

- A convenience sample is a sample from a population where we collect data on subjects because they're easy-to-find.
- **(4) Using your convenience sample, write and run code creating a `bargraph` for the number of people in each `grade`.**
    - **(5) Do you think the distribution of `grade` for your sample would look similar when compared to the whole `cdc` data?**
    - **(6) Which groups of people do you think are over- or under-represented in your convenience sample? Why?**
- **(7) Write and run code creating a `bargraph` for `grade` using the `cdc` data.**
    - **(8) Compare the distributions of the `cdc` data and your convenience sample and write down how they differ.**


## Using randomness

- **(9) Fill in the blanks below to create a sample by randomly selecting 50 people in the `cdc` data, without replacement. Call this new sample `s2`:**

```{r, eval = FALSE, echo = TRUE}
___ <- sample(___, size = ___, replace = ___)
```

- **(10) Write a sentence that explains why you think the distribution of `grade` for this _random sample_ will look more or less similar to the distribution from the whole `cdc` data.**
- **(11) Write and run code creating a `bargraph` for `grade` based on this _random sample_ to check your prediction.**
    

## Increasing sample size

- **(12) Write and run code creating `bargraph`s for `grade` based on each of the following sample sizes: 10, 100, 1,000, 10,000.**
     - Compare each distribution to that of the population.
- **(13) How do the distributions change as the size of the sample increases? Why do you think this occurs?**
- **(14) Write and run code to `tally()` the proportion of `grade`s for your _convenience_ sample and all your _random_ samples.**
    - **(15) Which set of proportions looks most similar to the proportions of the population?**
    

## Lessons learned

- The proportion (or mean) from a _random_ sample might not always be closer to that of the true population when compared to a _convenience_ sample. 
- However, as sample sizes get larger: 
    - _Random_ samples will tend to be better estimates for the population.
    -  With _convenience_ samples, this might not be the case.
- **(16) Write down a reason why estimates based on _convenience_ samples might not improve even as sample size increases.**