---
title: "All About Distributions"
author: "Lab 2A"
date: 'Directions: Follow along with the slides, completing the questions in <font color=#25679E><b>blue</b></font> on your computer, and answering the questions in **red** in your journal.  
  
  <br>
  <br>
  Space, Click, Right Arrow or swipe left to move to the next slide.'
output:
  slidy_presentation:
    css: ../../IDSLabCSSRev.css
    mathjax: ../../extras/mathjax-local/MathJax.js
  word_document:
  ioslides_presentation: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

## In the beginning ...

- Most of the labs thus far have covered how to visualize, summarize, and manipulate data.
    - We used visualizations to explore how your class spends their time.
    - We also learned how to clean data to prepare it for analyzing.
- Starting with this lab, we'll learn to use `R` to answer statistical investigative questions that can be answered by calculating the mean, median and MAD.


## How to talk about data

- When we make plots of our data, we usually want to know:
  - Where is the _bulk_ of the data?
  - Where is the data more _sparse_, or _thin_?
  - What values are _typical_?
  - How much does the data _vary_?
- To answer these questions, we want to look at the _distribution_ of our data.
    - We describe _distributions_ by talking about where the _center_ of the data are, how _spread_ out the data are, and what sort of _shape_ the data has.


## Let's begin!

- _Export_, _upload_ and _import_ your class's _Personality Color_ data.
    - Name your data `colors` when you load it.
- Before analyzing a new dataset, it's often helpful to get familiar with it. So:
    - **(1) Write down the `names` of the 4 variables that contain the point-totals, or _scores_, for each personality color.**
    - **(2) Write down the `names` of the variables that tell us an observation's introvert/extrovert designation and whether they are involved in _sports_.**
    - **(3) How many variables are in the dataset?**
    - **(4) How many observations are in the dataset?**


## Estimating centers

- **(5) Write and run code creating a `dotPlot` of the scores for _your_ predominant color.**
    - Pro-tip: If the `dotPlot` comes out looking wonky, include the `nint` and `cex` options.
- Based on your `dotPlot`:
    - **(6) Which values came up the most frequently? About how many people in  your class had a score similar to yours?**
    - **(7) What, would you say, was a _typical_ score for a person in your class for _your_ predominant color? How does your own score for this color compare?**


## Means and medians

- _Means_ and _medians_ are usually good ways to describe the _typical_ value of our data.
- **(8) Fill in the blank to calculate the `mean` value of _your_ predominant color score:**

```{r, eval = FALSE, echo = TRUE}
mean(~____, data = colors)
```

- **(9) Write and run a similar line of code to calculate the `median` value of _your_ predominant color.**
    - **(10) Are the `mean` and `median` roughly the same? If not, use the `dotPlot` you made in the last slide to describe why.**

## Estimating Spread

- Now that we know how to describe our data's _typical_ value we might also like to describe how closely the rest of the data are to this _typical_ value. 
    - We often refer to this as the <b>variability</b> of the data. 
    - Variability is seen in a `histogram` or `dotPlot` as the horizontal *spread*.
    
- **(11) Write and run code re-creating a `dotPlot` of the scores for _your_ predominant color and then write and run the code below filling in the blank with the name of _your_ predominant color:**

```{r, eval = FALSE, echo = TRUE}
add_line(vline = mean(~____, data = colors))
```

- **(12) Look at the spread of the scores from the mean score then complete the sentence below:**
    - **Data points in my plot will usually fall within ____ units of the center.**

## Mean Absolute Deviation

- The <b>mean absolute deviation</b> finds how far away, on average, the data are from the mean.
    - We often write _mean absolute deviation_ as _MAD_.
- **(13) Write and run code calculating the MAD of _your_ predominant color by filling in the blank:**

```{r, eval = FALSE, echo = TRUE}
MAD(~_____, data = colors)
```

- **(14) How close was your estimate of the spread for _your_ predominant color (from the previous slide) to the actual value?**

## Comparing introverts/extroverts

- Do introverts and extroverts differ in their typical scores for _your_ predominant color?
  - Answer this investigative question using a `dotPlot` and numerical summaries.
- **(15) Write and run code making a `dotPlot` of _your_ predominant color again; but this time, facet the plot by the introvert/extrovert variable. Include the `layout` option to stack the plots as well as the `nint` and `cex` options.**
- **(16) Describe the shape of the distribution of scores for the extroverts. Do the same for the introverts.**
- **(17) Using similar syntax to how you facet plots, write and run code _calculating_ either the `mean` or `median` to describe the _center_ of _your_ predominant color for introverts and extroverts.**
- **(18) Do introverts and extroverts differ in their typical scores for _your_ predominant color?**
- **(19) Based on the MAD, which group (introverts or extroverts) has more variability for _your_ predominant color's scores?**

## On your own

- Do introverts and extroverts in your class differ in their color scores? 
    - <span class="blue">Perform an analysis that produces _numerical summaries_ and _graphs_.</span>
    - **(20) Then, write a few sentences that address this statistical investigative question and considers the _shape_, _center_ and _spread_ of the distributions of the graphs you create.**