---
title: "Bootstrapping Lab"
author: "ThinkData Ed"
date: 'Directions: Follow along with the slides.  
  
  <br>
  <br>
  Space, Click, Right Arrow or swipe left to move to the next slide.'
output: 
  slidy_presentation: 
    css: ../../IDSLabCSSRev.css
    mathjax: ../../extras/mathjax-local/MathJax.js
---

## Introduction

**Research Question:** What is the mean age of people older than 15 living in the United States?

Today we'll use **bootstrapping** to create confidence intervals around our estimate.

**Key Concept:** A bootstrap sample is a random sample taken WITH replacement from our original data, using the same sample size.

## Load and Explore the Data
We will be using data from the American Time Use Survey. Run the following codes to load and _View_ the data.
```{r, echo = TRUE, eval = FALSE}
data(atus)
View(atus)

```

Now let's determine an estimate for the average age of people older than 15 living in the US:
```{r, echo = TRUE, eval = FALSE}
mean(~age, data = atus)
```

## Let's Create One Bootstrap Sample

Let's see what one bootstrapped sample looks like. To begin, we will set a seed for reproducibility:

```{r, echo = TRUE, eval = FALSE}
set.seed(123)
```

First, we will randomly pick which observations we want. We do this by sampling the numbers 1-8,136 (the size of the dataset) WITH replacement. 

```{r, echo = TRUE, eval = FALSE}

bs_rows <- sample(1:8136, size = 8136, replace = TRUE)
```

Now that we have which observations to sample, and again we want there to be some duplicated observations, we’ll use _slice()_ to use these rows to create a new set of data.
```{r, echo = TRUE, eval = FALSE}
bs_atus <- slice(atus, bs_rows)
```

Now let's find the mean of this sample:
```{r, echo = TRUE, eval = FALSE}
mean(~age, data = bs_atus)
```

## Make a Bootstrap Function
To make the process we did simple, let's create a function. 
```{r, echo = TRUE, eval = FALSE}
bs_func <- function() {
    bs_rows <- sample(1:8136, size = 8136, replace = TRUE)
    bs_atus <- slice(atus, bs_rows)
    mean(~age, data = bs_atus)
}

```

Now test it out to see. Try running it twice to see that you'll get multiple outputs.
```{r, echo = TRUE, eval = FALSE}
bs_func()
```

## Generate 2000 Bootstrap Samples

Let's run _bs_func()_ 2000 times. We will use the _do()_ function:
```{r, echo = TRUE, eval = FALSE}
bs_means <- do(2000)* bs_func()
```

Now let's visualize our bootstrapped means:
```{r, echo = TRUE, eval = FALSE}
histogram(~bs_func, data = bs_means,
          main = "Distribution of 2000 Bootstrap Sample Means")
```

Run the following for the summary statistics.
```{r, echo = TRUE, eval = FALSE}
favstats(~bs_func, data = bs_means)
```

## Create a 90% Confidence Interval
We're ready to make our confidence interval now! If we want to find a 90% confidence interval, we can look at the middle 90% of our data. To do this, we will find the 5% quantile and the 95% quantile. 

To find the lower bound, we will run:
```{r, echo = TRUE, eval = FALSE}
quantile(~bs_func, data = bs_means, p = 0.05)
```

For your upper bound, run:
```{r, echo = TRUE, eval = FALSE}
quantile(~bs_func, data = bs_means, p = 0.95)
```

Our confidence interval can be written with the format (lower bound, upper bound).

## Margin of Error
Margin of error is another way we can talk about how sure we are about our results. For example:

Instead of, 
  *"I think the mean age is between 52 and 53,"* we can say, 
  *"I think the mean age is 52.5, give or take 0.5 years."*

To find our margin of error, we can find the length of our interval and divide by two.

We'll store our 5% quantile as our lower bound:
```{r, echo = TRUE, eval = FALSE}
lower_bound<-quantile(~bs_func, data = bs_means, p = 0.05)
```

Similarly, let's store the 95% quantile as our upper bound:
```{r, echo = TRUE, eval = FALSE}
upper_bound<-quantile(~bs_func, data = bs_means, p = 0.95)
```

Now, we can find their difference and divide by two.
```{r, echo = TRUE, eval = FALSE}
(upper_bound - lower_bound)/2
```

That is our margin of error.

## What about a 95% Confidence Interval?

We'll simply adjust our quantiles! Our lower bound now is at 2.5%:
```{r, echo = TRUE, eval = FALSE}
quantile(~bs_func, data = bs_means, p = 0.025)
```

Our upper found will be at 97.5%.
```{r, echo = TRUE, eval = FALSE}
quantile(~bs_func, data = bs_means, p = 0.975)
```

Notice, the interval is wider now because we were willing to gain a little confidence for a wider interval. 

And a 95% confidence interval means that, if we were to take an infinite number of samples, our confidence interval would contain the true population mean 95% of the time.

## Summary

You've now learned how to:

  - Create bootstrap samples by sampling WITH replacement
   
  - Automate the process using functions
  
  - Generate many bootstrap samples to understand sampling variability
  
  - Calculate confidence intervals using quantiles
    
  - Find margin of error.

**Key Takeaway:** A 95% confidence interval means that if we repeated this process many times, about 95% of our intervals would contain the true population mean.